Analyzing X-ray images in Python with scikit-image

### 2016-08-21 skimage_tomo.tex

    \texttt{scikit-image} is an open-source image processing toolkit for
    the Python language that supports a large variety of file formats
    and is compatible with 2-D and 3-D images. The toolkit exposes a
    simple programming interface for functions that perform various
    operations on images, grouped thematically according to their
    purpose, e.g. image restoration,
    segmentation, measurements, etc. \texttt{scikit-image} users benefit
    from a rich scientific Python ecosystem that contains many
    powerful libraries for tasks such as visualization or machine learning.
    \texttt{scikit-image} provides a gentle
    learning curve, versatile image processing capabilities, and the
    scalable performances required for the high-throughput analysis of
    X-ray imaging data.

@@

Several software applications and libraries are available to synchrotron
users to process their images. ImageJ \citep{Abramoff2004, Schneider2012}
and its distribution Fiji \citep{Schindelin2012} is a popular tool for
2-D and 3-D images, thanks to its intuitive menus and graphical tools,
and the wealth of plugins contributed by a vivid
community \citep{Schindelin2015}. For 3-D images, commercial tools such
as Avizo 3D software (TM) are appreciated for an intuitive graphical
pipeline and advanced 3D visualization. Alternatively, the use of a
programming language gives finer control and more possibilities, provided
classical processing algorithms can be called from libraries -- thereby
limiting the complexity of the programming task and the risk of bugs. Matlab (TM) and its
image processing toolbox are popular in the academic community of
computer vision and image processing. The Python language is widely used
in the scientific world and in synchrotron facilities. As a
general-purpose language, Python is used in synchrotrons to control
device servers \citep{pytango}, to access raw data of X-ray
detectors \citep{Knudsen2013}, to reconstruct tomography volumes from
radiographs \citep{Gursoy2014, Mirone2014}, and in data
processing packages for macromolecular cristallography \citep{Adams2010},
azimuthal integration of diffraction data \citep{Ashiotis2015} or
fluorescence analysis \citep{pymca}.

@@

\texttt{scikit-image} has a very simple Application Programming Interface
(API), based almost exclusively on
functions. Most functions take an image (\emph{i.e} a multi-dimensional
array) as input parameter:
\begin{lstlisting}
>>> from skimage import filters
>>> im_gaussian = filters.gaussian(im)
\end{lstlisting}

@@

Most functions take both 2-D and 3-D images as arguments, so that
the same functions can be used to process tomography and microscopy
images. The rest raise an error
when passed a 3-D argument:
\begin{lstlisting}
>>> filters.prewitt(im)
[...]
ValueError: The parameter `image` must be a 2-dimensional array
\end{lstlisting}

@@

While several interpreters are available to execute Python
instructions and scripts interactively, the most popular in the scientific world is
IPython \citep{Perez2007, Rossant2015}. IPython is an
advanced interpreter, which integrates syntax highlighting, text auto-completion,
a debugger,
introspection and profiling methods, and online help. Several
Integrated Development Environments (IDEs) include IPython,
together with other components such as a text editor. Notable examples
include Spyder (Fig.~\ref{fig:spyder}), PyCharm, and Visual Studio Code.

@@

The Jupyter notebook \citep{Kluyver2016} is a web application that grew
out of the IPython project. Jupyter notebooks provide an interactive development
environment within a web browser, where live code can be enriched
by explanatory text, equations and visualizations (FIGURE).
(jni: FURTHER TEXT ABOUT IMPACT OF JUPYTER NOTEBOOKS)

\paragraph{NumPy arrays} are the cornerstone of the Scientific Python
ecosystem, and of \texttt{scikit-image} operations in particular.
Cropping or downsampling an image, or retrieving pixels corresponding to
a given label in a segmentation are all NumPy ``one-liners''. To
illustrate the compactness of NumPy code, consider modifying pixel values below a
threshold.  This operation can be written as
\begin{lstlisting}
im[im < 0.5] = 0
\end{lstlisting}
exploiting the possibility of indexing arrays with boolean arrays, also called
\emph{masking}. Also, \texttt{NumPy} uses memory sparingly and avoids
making new copies of arrays whenever possible, which is a important
requirement when dealing with the gigabyte-sized images of tomography.
For example, cropping a subvolume as follows does not create a copy of the
original array
\begin{lstlisting}
sub_volume = im[100:-100, 100:-100, 100:-100]
\end{lstlisting}
but instead refers to the correct memory offsets in the original.

@@

\texttt{scikit-image} is only one Python module that can be used for data
processing, among many others. A very popular module is
\texttt{scikit-learn} \citep{Pedregosa2011}, a Python module for machine
learning using \texttt{NumPy} arrays. Local features of an image
(such as local statistics of gray levels, or geometric points of
interest) or features of segmented objects (e.g. geometrical and
intensity characteristics of segmented particles) can be extracted with
functions from \texttt{skimage.feature} (see Fig.~\ref{fig:ecosystem}). It
is then possible to use a \emph{classification} algorithm from
\texttt{scikit-learn} to label pixels (a segmentation task)
or to classify whole images or objects that have already been segmented.
The interoperability between these packages is ensured by the common use
of \texttt{NumPy} arrays, so that the tasks described above are
implemented with a few lines of code.

The modularity of the Scientific Python ecosystem may be
confusing at first sight. Nevertheless, core modules of this ecosystem
are almost perfectly compatible, thanks to the shared use of NumPy arrays
and common development practices (although they are developed in parallel
by different teams). Several integrated distributions, such as Anaconda
or Canopy, bundle together the most common packages, including
\texttt{scikit-image}.

@@

\begin{figure*}
    \centerline{\includegraphics[width=0.99\textwidth]{tomo_gallery}}
    \caption{\csentence{Typical image processing operations with
	\texttt{scikit-image}.} From left to right and top to bottom.
	Data are syntheric, unless stated otherwise.
	\textbf{Filtering -} Top: non-local means denoising of an image
	with a fine-grained texture, acquired by \emph{in situ}
	synchrotron microtomography during
	glass melting \citep{Gouillart2012}. Bottom: total-variation
	denoising of an image with two phases, corresponding to
	phase-separating silicate melts observed by \emph{in situ
	tomography} \citep{Bouttes2015}.
	\textbf{Feature extraction -} Top: Hubble deep field (NASA,
	public domain), blob detection using the
	Laplacian of Gaussian method. Bottom: ridge detection using the
	leading value of the Hessian matrix. 
	\textbf{Segmentation - } Top: super-pixel segmentation
	of a CT slice of the human head \citep{tomo_wikipedia}, using
	Felzenszwalb's algorithm \citep{Felzenszwalb2004}. Bottom: random
	walker segmentation (right) of noisy image (top-left corner), using
	histogram-determined markers (bottom-left corner).
	\textbf{Measures -} Top: visualization of local diameter
	(color-coded on the skeleton curve) of an
	interconnected phase (represented in violet).  Bottom: particles color-coded according
	to their extent.
\label{fig:tomo_gallery}}
\end{figure*}

@@

Segmentation of regions of interest can be achieved using one of various
strategies, depending on the characteristics of the image. Images with a
clear contrast between regions can be segmented automatically thanks to
several thresholding algorithms, including an adaptive local thresholding
algorithm aimed at images with strong contrast variations. Super-pixel
algorithms \citep{Felzenszwalb2004, Achanta2012} (see
Fig.~\ref{fig:tomo_gallery}) create an over-segmentation of images in
super-pixels, by grouping pixels that are close together both in color
and spatial distance. Region-growing algorithms, such as the morphological
watershed or the random walker \citep{Grady2006}, propagate the labels
of user-defined markers through the image (see
Fig.~\ref{fig:tomo_gallery}). The active contour
algorithm \citep{Kass1988} fits snake contours to features of the
image, such as edges or high-brightness regions.

@@

An easy solution consists in dividing an image into smaller images (with
or without overlap, depending on the operation), and to apply the
same operation on the different sub-images, on different cores. Creating overlapping chunks is easy with the dedicated function \texttt{view\_as\_windows} (or \texttt{view\_as\_blocks} for contiguous non-overlapping chunks):
\begin{lstlisting}[language=Python]
>>> from skimage import util, data
>>> im = data.camera()
>>> im.shape
(512, 512)
>>> # chunks of size 120 with 8-pixel overlap
>>> chunks = util.view_as_windows(im, (128, 128), step=120)
>>> chunks.shape
(4, 4, 128, 128)
\end{lstlisting}
The \texttt{joblib} module enables easy parallel processing. Looping over the different blocks, and dispatching the computation over several cores, is realized with the following syntax:  
\begin{lstlisting}
>>> from joblib import Parallel, delayed
>>> filtered_chunks = Parallel(n_jobs=4)(delayed(filters.gaussian)(chunks[i, j]) for i in range(4) for j in range(4))
\end{lstlisting}
\texttt{scikit-image} also offers experimental support for a more integrated parallel processing pipeline, thanks to the \texttt{dask} \citep{dask} module:
\begin{lstlisting}
filtered_im = util.apply_parallel(filters.gaussian, im, depth=8)
\end{lstlisting}
The size of chunks is determined automatically from the number of available cpus, or can be specified by the user.

Another situation that often arises is that, while prototyping a
workflow, scripts (containing the image processing pipeline) are run
several times to experiment with parameters. The \texttt{joblib}
module provides a caching mechanism that avoids the repetition of function
calls if their arguments have not changed:
\begin{lstlisting}
>>> from skimage import filters, data
>>> from joblib import Memory
>>> mem = Memory(cachedir='/tmp/joblib')
>>> median = mem.cache(filters.median)
>>> im = data.camera()
>>> filtered_im = median(im, np.ones((3, 3)))
__________________________
[Memory] Calling skimage.filters.rank.generic.median...
median(array([[156, ..., 152],
       ..., 
       [121, ..., 111]], dtype=uint8), array([[ 1.,  1.,  1.],
       [ 1.,  1.,  1.],
       [ 1.,  1.,  1.]]))
__________________________median - 0.0s, 0.0min
>>> filtered_again = median(im, np.ones((3, 3)))
>>> # the above call did not trigger another evaluation
\end{lstlisting}

Finally, let us note that the biggest performance improvements often come from
the improvement of algorithms (as opposed to computing architectures only). For
example, non-local means denoising \citep{Buades2005} is a costly operation,
since it requires several nested loops, on all pixels and on neighboring
patches to be compared with the pixel-centered patch. Thanks to the
implementation of a more recent algorithm \citep{Darbon2008} that modifies the
internal organization of loops, it was possible to improve execution
time by a factor of roughly ten times.
The large size of the \texttt{scikit-image} community makes it likely for
algorithmic improvements to be discussed regularly.

@@

The quality of software documentation is (perhaps especially)
important in software aimed at scientists.
\texttt{scikit-image} users have access to several kinds of
documentation. Functions of the API are all
documented using the same standard \citep{Pawlik2015} as other major
Scientific Python packages. The standard include a description of all
input and output variables and their data types, together with
explanations of what each function does and how to use it. Function
documentation is accessed either online, or within the development
environment (IPython, Spyder, Jupyter Notebook...).

\begin{figure*}
    \centerline{\includegraphics[width=0.99\textwidth]{figure_gallery}}
\caption{\csentence{Gallery of examples of \texttt{scikit-image}.}
 The gallery of examples consists of an array of thumbnails (left part), which link to example webpages, each centerered on a specific image processing task. Each webpage includes Python code generating a figure, the figure itself, as well as a short tutorial explaining the image processing operations and the code. \label{fig:gallery}}
\end{figure*}

In addition, a graphical gallery of examples
(\nolinkurl{http://scikit-image.org/docs/dev/auto_examples/}),
part of which is displayed in Fig.~\ref{fig:gallery},
showcases graphical examples of image processing
operations.  The examples are organized as an array of thumbnails with a
short title (see Fig.~\ref{fig:gallery} left). These thumbnails link to
the webpage of the corresponding example, which features a mini-tutorial
on the image processing method, the code needed to run the example and
the figure generated by the example. Since the graphical gallery is an
efficient way to inform users about the features of
\texttt{scikit-image}, every new feature integrated in the package must
include an example for the gallery.
Longer tutorials and a more narrative documentation is available as well
in the online User Guide of \texttt{scikit-image}. The User Guide explains in
particular "big picture", foundational aspects of \texttt{scikit-image}, such as its
use of \texttt{NumPy} arrays as images, or how the package interacts with
other parts of the scientific Python ecosystem.

@@

Estimating the number of active users of an open-source package is a
difficult task. Download statistics, for example, largely
overestimate the number of active users, all the more if the package is
bundled with others in a software distribution, like Anaconda or
Canopy. A view closer to reality can be obtained by analyzing the
statistics of visits of the online help, available on the project
website.  As of the first half of 2016, 20000 unique visitors visited the
\texttt{scikit-image} website every month at \url{http://scikit-image.org/}. These
users come from 80 countries, located almost equally in Europe, America
and Asia (Oceania and Africa representing a much smaller fraction of
visitors).

@@

\texttt{scikit-image} is developed by a diverse team of volunteers, with
more than 170 developers having contributed to the package. The large
number of developers and users is a key element of the sustainability of
\texttt{scikit-image}. The
development process takes place on the GitHub platform
\url{https://github.com/scikit-image/scikit-image}, where users and
developers propose and discuss new contributions, report bugs or submit ideas
for improvements.
A release cycle of one or two releases every year
ensures that new features are propagated to users on a regular basis.

@@

\texttt{scikit-image} offers a wide variety of image processing
algorithms, using a simple interface natively compatible with 2D and 3D
images. It is well integrated into the Scientific Python ecosystem, so
that it interfaces well with visualization libraries and other data
processing packages. \texttt{scikit-image} has seen tremendous growth
since its creation in 2009, both in terms of users and proposed features.
In addition to the growing number of scientific teams that use
\texttt{scikit-image} for processing images of various X-ray modalities,
domain-specific tools are now using \texttt{scikit-image} as a dependency
to build upon. Examples include \texttt{tomopy} \citep{Gursoy2014} for
tomographic reconstruction, or \texttt{DIOPTAS} \citep{Prescher2015} for
the reduction and exploration of X-ray diffraction data. It is likely
that more application-specific software will benefit by depending on
\texttt{scikit-image} in the future, since \texttt{scikit-image} strives
to be domain-agnostic and to keep the function interface stable. On the end-user
side, future work includes better integration of parallel processing
capabilities, completion of full 3-D compatibility, an enriched narrative
documentation, speed enhancements, and expansion of the set of supported algorithms.

>>>>

    \texttt{scikit-image} is an open-source image processing toolkit for
    the Python language that supports a large variety of file formats and
    is compatible with 2-D and 3-D images. The toolkit exposes a simple
    programming interface, with thematic modules grouping functions
    according to their purpose, such as image restoration, segmentation,
    and measurements. \texttt{scikit-image} users benefit from a rich
    scientific Python ecosystem that contains many powerful libraries for
    tasks such as visualization or machine learning.
    \texttt{scikit-image} combines a gentle learning curve, versatile
    image processing capabilities, and the scalable performances required
    for the high-throughput analysis of X-ray imaging data.

@@

Several software applications and libraries are available to synchrotron
users to process their images. ImageJ \citep{Abramoff2004, Schneider2012}
and its distribution Fiji \citep{Schindelin2012} is a popular tool for
2-D and 3-D images, thanks to its intuitive menus and graphical tools,
and the wealth of plugins contributed by a vivid
community \citep{Schindelin2015}. For 3-D images, commercial tools such
as Avizo 3D software (TM) are appreciated for an intuitive graphical
pipeline and advanced 3D visualization. Alternatively, the use of a
programming language gives finer control and more possibilities, provided
classical processing algorithms can be called from libraries -- thereby
limiting the complexity of the programming task and the risk of bugs. Matlab (TM) and its
image processing toolbox are popular in the academic community of
computer vision and image processing. The Python language is widely used
in the scientific world and in synchrotron facilities. As a
general-purpose language, Python is used in synchrotrons to control
device servers \citep{pytango}, to access raw data of X-ray
detectors \citep{Knudsen2013}, to reconstruct tomography volumes from
radiographs \citep{Gursoy2014, Mirone2014}, and in data
processing packages for macromolecular cristallography \citep{Adams2010},
azimuthal integration of diffraction data \citep{Ashiotis2015}, or
fluorescence analysis \citep{pymca}.

@@

\texttt{scikit-image} has a simple Application Programming Interface
(API), based almost exclusively on
functions. Most functions take an image (\emph{i.e} a multi-dimensional
array) as input parameter:
\begin{lstlisting}
>>> from skimage import filters
>>> im_gaussian = filters.gaussian(im)
\end{lstlisting}

@@

Most functions transparently handle 2-D, 3-D, or even higher-dimensional
images as arguments, so
the same functions can be used to process tomography, microscopy, or natural
images. The rest raise an error when passed a 3-D argument:
\begin{lstlisting}
>>> filters.prewitt(im)
[...]
ValueError: The parameter `image` must be a 2-dimensional array
\end{lstlisting}
However, the proportion of functions supporting 3D images is always increasing,
thanks to the many contributors to the library.

@@

While several interpreters are available to execute Python instructions
and scripts interactively, the most popular in the scientific world is
IPython \citep{Perez2007, Rossant2015}. IPython is an advanced
interpreter, which integrates syntax highlighting, text auto-completion,
a debugger, introspection and profiling methods, and online help. Several
Integrated Development Environments (IDEs) come bundled with IPython,
together with other components such as a text editor. Notable examples
include Spyder (Fig.~\ref{fig:spyder}), PyCharm, and Visual Studio Code.

@@

The Jupyter notebook \citep{Kluyver2016} is a web application that grew
out of the IPython project. Jupyter notebooks provide an interactive development
environment within a web browser, where live code can be enriched
by explanatory text, equations and visualizations (FIGURE).
As of July 2016, more than 500,000 Jupyter notebooks were posted on GitHub,
demonstrating their wide adoption by the community as workflow-sharing
tools (\url{http://archive.ipython.org/media/SciPy2016JupyterLab.pdf}).

\paragraph{NumPy arrays} are the cornerstone of the Scientific Python
ecosystem, and of \texttt{scikit-image} operations in particular.
Cropping or downsampling an image, or retrieving pixels corresponding to
a given label in a segmentation are all NumPy ``one-liners''. To
illustrate the compactness of NumPy code, consider modifying pixel values below a
threshold.  This operation can be written as
\begin{lstlisting}
im[im < 0.5] = 0
\end{lstlisting}
exploiting the ability to index arrays with boolean arrays, also called
\emph{masking}. \texttt{NumPy} uses memory sparingly and avoids
making new copies of arrays whenever possible, an important
requirement when dealing with the gigabyte-sized images of tomography.
For example, cropping a subvolume as follows does not create a copy of the
original array
\begin{lstlisting}
sub_volume = im[100:-100, 100:-100, 100:-100]
\end{lstlisting}
but instead refers to the correct memory offsets in the original.

@@

\texttt{scikit-image} is only one Python module that can be used for data
processing, among many others. A very popular module is
\texttt{scikit-learn} \citep{Pedregosa2011}, a Python module for machine
learning using \texttt{NumPy} arrays. Local features of an image
(such as local statistics of gray levels, or geometric points of
interest) or features of segmented objects (e.g. geometrical and
intensity characteristics of segmented particles) can be extracted with
functions from \texttt{skimage.feature} (see Fig.~\ref{fig:ecosystem}). It
is then possible to use a \emph{classification} algorithm from
\texttt{scikit-learn} to label pixels (a segmentation task)
or to classify whole images or objects that have already been segmented.
The near-universal use of \texttt{NumPy} arrays ensures the interoperability
between these packages, so that just a few lines of code are sufficient
to create these sophisticated workflows.

The modularity of the Scientific Python ecosystem may be
confusing at first sight, but the core modules of this ecosystem
are almost perfectly compatible, thanks to the shared use of NumPy arrays
and common development practices (although they are developed in parallel
by different teams). Several ``distributions'', such as Anaconda
or Canopy, bundle together the most popular libraries, including
\texttt{scikit-image}.

@@

\begin{figure*}
    \centerline{\includegraphics[width=0.99\textwidth]{tomo_gallery}}
    \caption{\csentence{Typical image processing operations with
	\texttt{scikit-image}.} From left to right and top to bottom.
	Data are synthetic, unless stated otherwise.
	\textbf{Filtering -} Top: non-local means denoising of an image
	with a fine-grained texture, acquired by \emph{in situ}
	synchrotron microtomography during
	glass melting \citep{Gouillart2012}. Bottom: total-variation
	denoising of an image with two phases, corresponding to
	phase-separating silicate melts observed by \emph{in situ
	tomography} \citep{Bouttes2015}.
	\textbf{Feature extraction -} Top: Hubble deep field (NASA,
	public domain), blob detection using the
	Laplacian of Gaussian method. Bottom: ridge detection using the
	leading eigenvalue of the Hessian matrix.
	\textbf{Segmentation - } Top: super-pixel segmentation
	of a CT slice of the human head \citep{tomo_wikipedia}, using
	Felzenszwalb's algorithm \citep{Felzenszwalb2004}. Bottom: random
	walker segmentation (right) of noisy image (top-left corner), using
	histogram-determined markers (bottom-left corner).
	\textbf{Measures -} Top: visualization of local diameter
	(color-coded on the skeleton curve) of an
	interconnected phase (represented in violet).  Bottom: particles color-coded according
	to their extent.
\label{fig:tomo_gallery}}
\end{figure*}

@@

Segmentation of regions of interest can be achieved using one of various
strategies, depending on the characteristics of the image. Images with a
clear contrast between regions can be segmented automatically thanks to
several thresholding algorithms, including an adaptive local thresholding
algorithm aimed at images with contrast variations. Super-pixel
algorithms \citep{Felzenszwalb2004, Achanta2012} create an
over-segmentation of images in super-pixels, by grouping pixels that are
close together both in color- and spatial distance (see
Fig.~\ref{fig:tomo_gallery}). Region-growing algorithms, such as the
morphological watershed or the random walker \citep{Grady2006}, propagate
the labels of user-defined markers through the image (see
Fig.~\ref{fig:tomo_gallery}). The active contour algorithm
\citep{Kass1988} fits snake contours to features of the image, such as
edges or high-brightness regions.

@@

An easy solution consists in dividing an image into smaller images (with
or without overlap, depending on the operation), and to apply the
same operation on the different sub-images, on different cores. Creating overlapping chunks is easy with the dedicated function \texttt{view\_as\_windows} (or \texttt{view\_as\_blocks} for contiguous non-overlapping chunks):
\begin{lstlisting}[language=Python]
>>> from skimage import util, data
>>> im = data.camera()
>>> im.shape
(512, 512)
>>> # chunks of size 120 with 8-pixel overlap
>>> chunks = util.view_as_windows(im, (128, 128), step=120)
>>> chunks.shape
(4, 4, 128, 128)
\end{lstlisting}
The \texttt{joblib} library enables easy parallel processing. Looping over the different blocks, and dispatching the computation over several cores, is realized with the following syntax:
\begin{lstlisting}
>>> from joblib import Parallel, delayed
>>> filtered_chunks = Parallel(n_jobs=4)(delayed(filters.gaussian)(chunks[i, j]) for i in range(4) for j in range(4))
\end{lstlisting}
\texttt{scikit-image} also offers experimental support for a more integrated parallel processing pipeline, thanks to the \texttt{dask} \citep{dask} module:
\begin{lstlisting}
filtered_im = util.apply_parallel(filters.gaussian, im, depth=8)
\end{lstlisting}
The size of chunks is determined automatically from the number of available cpus, or can be specified by the user.

\em{Caching} provides another tool to speed up data analysis. A situation
that often arises is that, while prototyping a workflow, scripts
(containing the image processing pipeline) are run several times to
experiment with parameters. \texttt{joblib} provides a caching mechanism
that avoids the repetition of function calls, if their arguments have not
changed:
\begin{lstlisting}
>>> from skimage import filters, data
>>> from joblib import Memory
>>> mem = Memory(cachedir='/tmp/joblib')
>>> median = mem.cache(filters.median)
>>> im = data.camera()
>>> filtered_im = median(im, np.ones((3, 3)))
__________________________
[Memory] Calling skimage.filters.rank.generic.median...
median(array([[156, ..., 152],
       ..., 
       [121, ..., 111]], dtype=uint8), array([[ 1.,  1.,  1.],
       [ 1.,  1.,  1.],
       [ 1.,  1.,  1.]]))
__________________________median - 0.0s, 0.0min
>>> filtered_again = median(im, np.ones((3, 3)))
>>> # the above call did not trigger another evaluation
\end{lstlisting}

Finally, we note that the biggest performance improvements often come from
the improvement of algorithms (as opposed to computing architectures only). For
example, non-local means denoising \citep{Buades2005} is a costly operation,
since it requires several nested loops, on all pixels and on neighboring
patches to be compared with the pixel-centered patch. Thanks to the
implementation of a more recent algorithm \citep{Darbon2008} that modifies the
internal organization of loops, it was possible to improve execution
time by a factor of roughly ten times.
The large size of the \texttt{scikit-image} community makes it likely for
algorithmic improvements to be discussed regularly.

@@

The quality of software documentation is (perhaps especially) important
in software aimed at scientists. \texttt{scikit-image} users have access
to several kinds of documentation. All functions are documented using the
NumPy documentation standard \citep{Pawlik2015}, which is universal
across all major Scientific Python packages. The standard include a
description of all input and output variables and their data types,
together with explanations of what each function does and how to use it.
Function documentation is accessible online or within the development
environment itself (IPython, Spyder, Jupyter Notebook...).

\begin{figure*}
    \centerline{\includegraphics[width=0.99\textwidth]{figure_gallery}}
\caption{\csentence{Gallery of examples of \texttt{scikit-image}.}
 The gallery of examples consists of an array of thumbnails (left), which link to example webpages, each centerered on a specific image processing task. Each webpage includes Python code generating a figure, the figure itself, and a short tutorial explaining the image processing operations and the code. \label{fig:gallery}}
\end{figure*}

In addition, a graphical gallery of examples
(\nolinkurl{http://scikit-image.org/docs/dev/auto_examples/}), part of
which is displayed in Fig.~\ref{fig:gallery}, showcases graphical
examples of common image processing operations.  The examples are
organized as an array of thumbnails with a short title (see
Fig.~\ref{fig:gallery} left). These thumbnails link to the webpage of the
corresponding example, which features a mini-tutorial on the image
processing method, the code needed to run the example and the figure
generated by the example. Since the graphical gallery is an efficient way
to inform users about the features of \texttt{scikit-image}, every new
feature integrated in the package must include an example for the
gallery. Longer tutorials and a more narrative documentation is available
as well in the online User Guide of \texttt{scikit-image}. The User Guide
explains in particular "big picture", foundational aspects of
\texttt{scikit-image}, such as its use of \texttt{NumPy} arrays as
images, or how the package interacts with other parts of the scientific
Python ecosystem.

@@

Estimating the number of active users of an open-source package is a
difficult task. Download statistics, for example, largely
overestimate the number of active users, all the more if the package is
bundled with others in a software distribution, such as Anaconda or
Canopy. A view closer to reality can be obtained by analyzing the
statistics of visits of the online help, available on the project
website.  As of the first half of 2016, 20000 unique visitors visited the
\texttt{scikit-image} website every month at \url{http://scikit-image.org/}. These
users come from 80 countries, located almost equally in Europe, America
and Asia (with Oceania and Africa representing a much smaller fraction of
visitors).

@@

\texttt{scikit-image} is developed by a diverse team of volunteers.
More than 170 individuals have contributed to the package. The large
number of developers and users is a key element of the sustainability of
\texttt{scikit-image}. The
development process takes place on GitHub
\url{https://github.com/scikit-image/scikit-image}, where users and
developers propose and discuss new contributions, report bugs or submit ideas
for improvements.
A release cycle of one or two releases every year
ensures that new features are propagated to users on a regular basis.

@@

\texttt{scikit-image} offers a wide variety of image processing
algorithms, using a simple interface natively compatible with 2D and 3D
images. It is well integrated into the Scientific Python ecosystem, so
that it interfaces well with visualization libraries and other data
processing packages. \texttt{scikit-image} has seen tremendous growth
since its creation in 2009, both in terms of users and proposed features.
In addition to the growing number of scientific teams that use
\texttt{scikit-image} for processing images of various X-ray modalities,
domain-specific tools are now using \texttt{scikit-image} as a dependency
to build upon. Examples include \texttt{tomopy} \citep{Gursoy2014} for
tomographic reconstruction, or \texttt{DIOPTAS} \citep{Prescher2015} for
the reduction and exploration of X-ray diffraction data. It is likely
that more application-specific software will benefit from depending on
\texttt{scikit-image} in the future, since \texttt{scikit-image} strives
to be domain-agnostic and to keep the function interface stable. On the end-user
side, future work includes better integration of parallel processing
capabilities, completion of full 3-D compatibility, an enriched narrative
documentation, speed enhancements, and expansion of the set of supported algorithms.
